---
title: "gwas_visualization"
output: html_document
date: "2025-01-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(qqman)
library(tidyr)
library(dplyr)
library(ggplot2)
library(manhattanly)
```

## Both Morphine and Fentanyl Usage Phenotype
```{r}
gwas_both <- read.table("phenos.BT.step2_both_morphine_fentanyl.txt", header =T, as.is = T, sep ='\t')
head(gwas_both)
```

Adding P column by inversing negativo logarithm with the base 10

```{r}
gwas_both <- gwas_both %>% mutate(P=(10^(-LOG10P)))

head(gwas_both)
```

```{r}
# Dimensions of the dataframe
dim(gwas_both)
```

```{r}
#Lowest P-value
min(gwas_both$P)
```

### Generating a Q-Q plot

```{r}
qq(gwas_both$P, main = "Q-Q plot for Both Fentanyl and Morphine Usage Phenotype Genome Association")
```
### Plotting a Manhattan Plot

We can use the manhattan() function form the {qqman} package to generate a manhattan plot.

```{r}
manhattan(gwas_both, chr="CHROM", bp ="GENPOS", snp="ID", p ="P", ylim=c(0,10), suggestiveline=FALSE, col="blue", main="Manhattan Plot for Both Fentanyl and Morphine Usage Phenotype Genome Association")
```

We can also use the {manhattaly} package to make an interactive manhattan plot

```{r}
manhattanly(gwas_both,chr="CHROM", bp ="GENPOS", snp="ID",annotation1 = "CHISQ", annotation2= "BETA", p ="P", suggestiveline=FALSE, col="blue")
```


Now let's do the same for the the other two phenotypes

## Only Morphine phenotype

```{r}
gwas_morphine <- read.table("phenos.BT.step2_only_morphine.txt", header =T, as.is = T, sep ='\t')
head(gwas_morphine)
```

Adding P column by inversing negativo logarithm with the base 10

```{r}
gwas_morphine <- gwas_morphine %>% mutate(P=(10^(-LOG10P)))

head(gwas_morphine)
```

```{r}
# Dimensions of the dataframe
dim(gwas_morphine)
```

```{r}
#Lowest P-value
min(gwas_morphine$P)
```

### Generating a Q-Q plot

```{r}
qq(gwas_morphine$P, main = "Q-Q plot for Only Morphine Usage Phenotype Genome Association")
```
### Plotting a Manhattan Plot

We can use the manhattan() function form the {qqman} package to generate a manhattan plot.

```{r}
manhattan(gwas_morphine, chr="CHROM", bp ="GENPOS", snp="ID", p ="P", ylim=c(0,10), suggestiveline=FALSE, col="blue", main="Manhattan Plot for Only Morphine Usage Phenotype Genome Association")
```

We can also use the {manhattaly} package to make an interactive manhattan plot

```{r}
manhattanly(gwas_morphine,chr="CHROM", bp ="GENPOS", snp="ID",annotation1 = "CHISQ", annotation2= "BETA", p ="P", suggestiveline=FALSE, col="blue")
```


## Only Fentanyl phenotype

```{r}
gwas_fentanyl <- read.table("phenos.BT.step2_only_fentanyl.txt", header =T, as.is = T, sep ='\t')
head(gwas_fentanyl)
```

Adding P column by inversing negativo logarithm with the base 10

```{r}
gwas_fentanyl <- gwas_fentanyl %>% mutate(P=(10^(-LOG10P)))

head(gwas_fentanyl)
```

```{r}
# Dimensions of the dataframe
dim(gwas_fentanyl)
```

```{r}
#Lowest P-value
min(gwas_fentanyl$P)
```

### Generating a Q-Q plot

```{r}
qq(gwas_fentanyl$P, main = "Q-Q plot for Only Fentanyl Usage Phenotype Genome Association")
```
### Plotting a Manhattan Plot

We can use the manhattan() function form the {qqman} package to generate a manhattan plot.

```{r}
manhattan(gwas_fentanyl, chr="CHROM", bp ="GENPOS", snp="ID", p ="P", ylim=c(0,10), suggestiveline=FALSE, col="blue", main="Manhattan Plot for Only Fentanyl Usage Case Genome Association")
```

We can also use the {manhattaly} package to make an interactive manhattan plot

```{r}
manhattanly(gwas_fentanyl,chr="CHROM", bp ="GENPOS", snp="ID",annotation1 = "CHISQ", annotation2= "BETA", p ="P", suggestiveline=FALSE, col="blue")
```
